Cell Decompositions for Rank Two Quiver Grassmannians

### 2019-06-09 Grassmannian_cells.tex
Notation changes in section 4

  For the reverse direction, assume $U\in X_{\bfi}$ is a $\CC^*$-fixed point represented by the matrix $M(U)$.
  Then, if~$v_{i_q}\in X_l$, the assumptions on $d$ imply $M(U)_{qr}=0$ unless $r_{l-1}+1\le r\le r_l$.
  That is, $M(U)$ has the shape of a block matrix representing the decomposition $U=\bigoplus_{l=1}^m U\cap X_l$.
\qed\end{proof}
The next step is to generalize this to quiver Grassmannians.
Let $Q$ be an acyclic quiver.
Choose a map $d:\hat Q_0\to\ZZ$ and fix a representation $X\in\rep Q$ which can be lifted to $\hat Q$.
We consider the decomposition $X_i=\bigoplus_{\chi\in A_Q} X_{(i,\chi)}$ and define a $\CC^*$-action on each $X_{(i,\chi)}$ via $t.x_{(i,\chi)}=t^{d(i,\chi)}x_{(i,\chi)}$ which is then extended linearly to each $X_i$.
Associated to each subspace $U_i$, there is a corresponding subspace $t.U_i$ for each $t\in\CC^*$.
In general, this does not induce a $\CC^*$-action on the quiver Grassmannians $\Gr_\bfe(X)$ since $t.U=(t.U_i)_{i\in Q_0}$ is not necessarily a subrepresentation of $X$ for every $U\in\Gr_\bfe(X)$.
Indeed, for this such an action must satisfy $X_\alpha(t.U_i)\subset t.U_j$ for every arrow $\alpha:i\to j$ of $Q$ and every $t\in\CC^*$. 
\begin{lemma}
  Let $X$ be a representation of $Q$ which can be lifted to $\hat Q$.
  Fix an integer $c_\alpha\in\ZZ$ for each~$\alpha\in Q_1$.
  Suppose $d:\hat Q_0\to\ZZ$ satisfies $d(j,\chi+e_\alpha)-d(i,\chi)=c_\alpha$ for each arrow $\alpha:i\to j$ of $Q$ and each $\chi\in A_Q$.
  Then the $\CC^*$-action on $X$ determined by $d$ induces a $\CC^*$-action on $\Gr_\bfe(X)$.
\end{lemma}
\begin{proof} 
  Fix $U\in\Gr_\bfe(X)$ and consider $u_i\in U_i$.
  Since $U$ is a subrepresentation, for an arrow $\alpha:i\to j$ of $Q$ we may write $X_\alpha(u_i)=u_j$ for some $u_j\in U_j$.

  As $X$ can be lifted to $\hat Q$, for any arrow $\alpha\in Q_1$ we can write $X_\alpha:X_i\to X_j$ as a block matrix consisting of linear maps $X_{(\alpha,\chi)}:X_{(i,\chi)}\to X_{(j,\chi+e_\alpha)}$ for $\chi\in A_Q$.
  Then writing $u_i=\sum_{\chi\in A_Q} u_{(i,\chi)}$ for some vectors $u_{(i,\chi)}\in X_{(i,\chi)}$, we have $X_\alpha(u_{(i,\chi)})=X_{(\alpha,\chi)}(u_{(i,\chi)})\in X_{(j,\chi+e_\alpha)}$, say $X_{(\alpha,\chi)}(u_{(i,\chi)})=u_{(j,\chi+e_\alpha)}$.
  It follows that $u_j=\sum_{\chi\in A_Q} u_{(j,\chi+e_\alpha)}$ and so
  \begin{align*}
    X_\alpha(t.u_i)
    &=\sum\limits_{\chi\in A_Q} t^{d(i,\chi)}X_{(\alpha,\chi)}(u_{(i,\chi)})=\sum\limits_{\chi\in A_Q} t^{d(i,\chi)}u_{(j,\chi+e_\alpha)}=t^{-c_\alpha}\sum\limits_{\chi\in A_Q} t.u_{(j,\chi+e_\alpha)}=t^{-c_\alpha} t.u_j.
  \end{align*}
  Therefore $X_\alpha(t.U_i)\subset t.U_j$ for every arrow $\alpha:i\to j$ of $Q$ and we obtain a $\CC^*$-action on $\Gr_\bfe(X)$.
\qed\end{proof}

@@

In the following, we say that $d:\supp(\hat X)\to\ZZ$ satisfies the \emph{degree condition} for $\hat X$ if it has the properties of Lemma~\ref{le:degree condition}.
\begin{theorem}
  \label{thm:torusfixedpoints}
  Let $X$ be a representation of $Q$ which can be lifted to a representation $\hat X$ of $\hat Q$ and choose $d:\supp(\hat X)\to\ZZ$ such that it satisfies the degree condition for $\hat X$.
  Then the $\CC^*$-action on $\bigoplus_{i\in Q_0} X_i$ determined by $t.x_{(i,\chi)}=t^{d(i,\chi)}x_{(i,\chi)}$ for $x_{(i,\chi)}\in X_{(i,\chi)}$ induces a $\CC^*$-action on $\Gr_\bfe^Q(X)$ such that
  \[\Gr^Q_\bfe(X)^{\CC^*}\cong\bigsqcup_{\hat\bfe} \Gr^{\hat Q}_{\hat\bfe}(\hat X),\]
  where $\hat\bfe$ runs through all dimension vectors compatible with $\bfe$.
\end{theorem}
\begin{proof}
  A representation $U\in\Gr_{\bfe}(X)$ is a $\CC^*$-fixed point if and only if $t.U=U$ for all $t\in\C^\ast$, i.e.\ $t.U_i=U_i$ for all $i\in Q_0$ and all $t\in\C^\ast$.
  Thus, apart from being a subrepresentation of $X$, each component $U_i$ is a fixed point of the induced $\CC^*$-actions on the usual Grassmannians of vector subspaces $\Gr_{\bfe_i}(X_i)$.
  By Lemma~\ref{le:usualGrass}, this holds precisely when we have a decomposition 
  \[U_i=\bigoplus_{\chi\in A_Q} U_i\cap X_{(i,\chi)}\]
  which is equivalent to $U$ being liftable to the universal abelian covering $\hat Q$.
\qed\end{proof}

@@

\begin{theorem}
  \label{thm:truncpp}
  Fix a dimension vector $\bfe$.
  The quiver Grassmannian $\Gr_\bfe(P_{m+1}^V)$ is smooth for each $V\in \Gr(\cH_m)$.
  Moreover, for $V,W\in \Gr_d(\cH_m)$, we have $\Gr_\bfe(P_{m+1}^V)\cong \Gr_\bfe(P_{m+1}^W)$.
\end{theorem}
\begin{proof}
  Let $g\in\GL_n(\CC)$.
  We first show that $\Gr_\bfe(g.P_{m+1}^V)=\Gr_\bfe(P_{m+1}^V)$.
  Assume that $P_{m+1}^V$ is given by the linear maps $M_{\alpha_i}$.
  Then $g.P_{m+1}^V$ is given by the matrices $(g.M)_{\alpha_i}=\sum\limits_{j=1}^n g_{ij}M_{\alpha_j}$.
  Let $(U_1,U_2)\in\Gr_\bfe(g.P_{m+1}^V)$.
  Then we have  
  \[M_{\alpha_i}(U_2)\subset U_1 \text{ for all } i=1,\ldots,n\Leftrightarrow \left(\sum\limits_{j=1}^n g_{ij}M_{\alpha_j}\right)(U_2)\subset U_1 \text{ for all } i=1,\ldots,n.\]
  Note that we have $g^{-1}.(g.P_{m+1}^V)=P_{m+1}^V$ which shows the non-obvious direction.

@@

  \noindent{\bf Claim.}
  For any admissible sequence $\bfI=(I_0,\ldots,I_k)$ with $k<m$ and any subset $J\subset I_0$, the quiver Grassmannian $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ admits a cell decomposition which is compatible with the unique map $\pi_{m+1}^{J,\bfI}:\tilde P_{m+1}^J\to\tilde P_{m+1}^\bfI$ from Lemma~\ref{le:truncated quotients} in the following sense:
  \begin{itemize}
    \item[($\dagger$)] For any $E\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ such that $\pi_{m+1}^{J,\bfI}(E)\neq 0$, we have $\pi_{m+1}^{J,\bfI}(E')\neq 0$ for all $E'\in C_E$, where $C_E$ is the affine cell which contains $E$.
  \end{itemize}
  \vs

>>>>

  For the reverse direction, assume $U\in X_{\bfi}$ is a $\CC^*$-fixed point represented by the matrix $M(U)$.
  Then, if~$v_{i_q}\in X_l$, the assumptions on $d$ imply $M(U)_{qr}=0$ unless $r_{l-1}+1\le r\le r_l$.
  That is, $M(U)$ has the shape of a block matrix representing the decomposition $U=\bigoplus_{l=1}^m U\cap X_l$.
\qed\end{proof}
The next step is to generalize this to quiver Grassmannians.
Let $Q$ be an acyclic quiver.
Choose a map $d:\hat Q_0\to\ZZ$ and fix a representation $X\in\rep Q$ which can be lifted to $\hat Q$.
We consider the decomposition $X_i=\bigoplus_{\chi\in A_Q} X_{(i,\chi)}$ and define a $\CC^*$-action on each $X_{(i,\chi)}$ via $t.x_{(i,\chi)}=t^{d(i,\chi)}x_{(i,\chi)}$ which is then extended linearly to each $X_i$.
Associated to each subspace $E_i$, there is a corresponding subspace $t.E_i$ for each $t\in\CC^*$.
In general, this does not induce a $\CC^*$-action on the quiver Grassmannians $\Gr_\bfe(X)$ since $t.E=(t.E_i)_{i\in Q_0}$ is not necessarily a subrepresentation of $X$ for every $E\in\Gr_\bfe(X)$.
Indeed, for this such an action must satisfy $X_\alpha(t.E_i)\subset t.E_j$ for every arrow $\alpha:i\to j$ of $Q$ and every $t\in\CC^*$. 
\begin{lemma}
  Let $X$ be a representation of $Q$ which can be lifted to $\hat Q$.
  Fix an integer $c_\alpha\in\ZZ$ for each~$\alpha\in Q_1$.
  Suppose $d:\hat Q_0\to\ZZ$ satisfies $d(j,\chi+e_\alpha)-d(i,\chi)=c_\alpha$ for each arrow $\alpha:i\to j$ of $Q$ and each $\chi\in A_Q$.
  Then the $\CC^*$-action on $X$ determined by $d$ induces a $\CC^*$-action on $\Gr_\bfe(X)$.
\end{lemma}
\begin{proof} 
  Fix $E\in\Gr_\bfe(X)$ and consider $e_i\in E_i$.
  Since $E$ is a subrepresentation, for an arrow $\alpha:i\to j$ of $Q$ we may write $X_\alpha(e_i)=e_j$ for some $e_j\in E_j$.

  As $X$ can be lifted to $\hat Q$, for any arrow $\alpha\in Q_1$ we can write $X_\alpha:X_i\to X_j$ as a block matrix consisting of linear maps $X_{(\alpha,\chi)}:X_{(i,\chi)}\to X_{(j,\chi+e_\alpha)}$ for $\chi\in A_Q$.
  Then writing $e_i=\sum_{\chi\in A_Q} e_{(i,\chi)}$ for some vectors $e_{(i,\chi)}\in X_{(i,\chi)}$, we have $X_\alpha(e_{(i,\chi)})=X_{(\alpha,\chi)}(e_{(i,\chi)})\in X_{(j,\chi+e_\alpha)}$, say $X_{(\alpha,\chi)}(e_{(i,\chi)})=e_{(j,\chi+e_\alpha)}$.
  It follows that $e_j=\sum_{\chi\in A_Q} e_{(j,\chi+e_\alpha)}$ and so
  \begin{align*}
    X_\alpha(t.e_i)
    &=\sum\limits_{\chi\in A_Q} t^{d(i,\chi)}X_{(\alpha,\chi)}(e_{(i,\chi)})=\sum\limits_{\chi\in A_Q} t^{d(i,\chi)}e_{(j,\chi+e_\alpha)}=t^{-c_\alpha}\sum\limits_{\chi\in A_Q} t.e_{(j,\chi+e_\alpha)}=t^{-c_\alpha} t.e_j.
  \end{align*}
  Therefore $X_\alpha(t.E_i)\subset t.E_j$ for every arrow $\alpha:i\to j$ of $Q$ and we obtain a $\CC^*$-action on $\Gr_\bfe(X)$.
\qed\end{proof}

@@

In the following, we say that $d:\supp(\hat X)\to\ZZ$ satisfies the \emph{degree condition} for $\hat X$ if it has the properties of Lemma~\ref{le:degree condition}.
\begin{theorem}
  \label{thm:torusfixedpoints}
  Let $X$ be a representation of $Q$ which can be lifted to a representation $\hat X$ of $\hat Q$ and choose $d:\supp(\hat X)\to\ZZ$ such that it satisfies the degree condition for $\hat X$.
  Then the $\CC^*$-action on $\bigoplus_{i\in Q_0} X_i$ determined by $t.x_{(i,\chi)}=t^{d(i,\chi)}x_{(i,\chi)}$ for $x_{(i,\chi)}\in X_{(i,\chi)}$ induces a $\CC^*$-action on $\Gr_\bfe^Q(X)$ such that
  \[\Gr^Q_\bfe(X)^{\CC^*}\cong\bigsqcup_{\hat\bfe} \Gr^{\hat Q}_{\hat\bfe}(\hat X),\]
  where $\hat\bfe$ runs through all dimension vectors compatible with $\bfe$.
\end{theorem}
\begin{proof}
  A representation $E\in\Gr_{\bfe}(X)$ is a $\CC^*$-fixed point if and only if $t.E=E$ for all $t\in\C^\ast$, i.e.\ $t.E_i=E_i$ for all $i\in Q_0$ and all $t\in\C^\ast$.
  Thus, apart from being a subrepresentation of $X$, each component $E_i$ is a fixed point of the induced $\CC^*$-actions on the usual Grassmannians of vector subspaces $\Gr_{\bfe_i}(X_i)$.
  By Lemma~\ref{le:usualGrass}, this holds precisely when we have a decomposition 
  \[E_i=\bigoplus_{\chi\in A_Q} E_i\cap X_{(i,\chi)}\]
  which is equivalent to $E$ being liftable to the universal abelian covering $\hat Q$.
\qed\end{proof}

@@

\begin{theorem}
  \label{thm:truncpp}
  Fix a dimension vector $\bfe$.
  The quiver Grassmannian $\Gr_\bfe(P_{m+1}^V)$ is smooth for each $V\in \Gr(\cH_m)$.
  Moreover, for $V,W\in \Gr_d(\cH_m)$, we have $\Gr_\bfe(P_{m+1}^V)\cong \Gr_\bfe(P_{m+1}^W)$.
\end{theorem}
\begin{proof}
  Let $g\in\GL_n(\CC)$.
  We first show that $\Gr_\bfe(g.P_{m+1}^V)=\Gr_\bfe(P_{m+1}^V)$.
  Assume that $P_{m+1}^V$ is given by the linear maps $M_{\alpha_i}$.
  Then $g.P_{m+1}^V$ is given by the matrices $(g.M)_{\alpha_i}=\sum\limits_{j=1}^n g_{ij}M_{\alpha_j}$.
  Let $(E_1,E_2)\in\Gr_\bfe(g.P_{m+1}^V)$.
  Then we have  
  \[M_{\alpha_i}(E_2)\subset E_1 \text{ for all } i=1,\ldots,n\Leftrightarrow \left(\sum\limits_{j=1}^n g_{ij}M_{\alpha_j}\right)(E_2)\subset E_1 \text{ for all } i=1,\ldots,n.\]
  Note that we have $g^{-1}.(g.P_{m+1}^V)=P_{m+1}^V$ which shows the non-obvious direction.

@@

  \noindent{\bf Claim.}
  For any admissible sequence $\bfi=(i_1,\ldots,i_k)$ with $k<m$ and any subsets $I,J\subsetneq\{1,\ldots,n\}$ with $i_k\in I$ and $i_1\notin J$, the quiver Grassmannian $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ admits a cell decomposition which is compatible with the unique map $\tilde\pi_{m+1}^{\ui,J,I}:\tilde P_{m+1}^J\to\tilde P_{m+1-k,\ui}^I$ from Lemma~\ref{le:truncated quotients} in the following sense:
  \begin{itemize}
    \item[($\dagger$)] For any $E\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ such that $\tilde\pi_{m+1}^{\ui,J,I}(E)\neq 0$, we have $\tilde\pi_{m+1}^{\ui,J,I}(E')\neq 0$ for all $E'\in C_E$, where $C_E$ is the affine cell which contains $E$.
  \end{itemize}
  \vs
